Beyond Vectors - Beginning with Data frames
===========================================

A data frame is a very important data type in R. It's pretty much the de facto data structure for most tabular data and what we use for statistics. Data frames can have additional attributes such as rownames() and colnames(). This can be useful for annotating data.

**Useful functions**  

head() - see first 5 rows

tail() - see last 5 rows

dim() - see dimensions

nrow() - number of rows

ncol() - number of columns

str() - structure of each column

names() - will list column names for a data.frame (or any object really).

Data frames are usually read in from file, but R comes with many practice datasets. We will use the iris dataset, famously used by Fisher in 1936 (http://en.wikipedia.org/wiki/Iris_flower_data_set)

```{r}
iris
```

**Try It!** A.
------------
1. View the dataset using head and tail
2. How many rows does the data have?
3. How many columns? What are the column names?
4. Using the str function, how many species are in the data?

Dataframe Syntax and subsetting
-------------------------------
R has many powerful subset operators and mastering them will allow you to easily perform complex operation on any kind of dataset. Allows you to manipulate data very succinctly.

While not technically accurate, dataframes are akin to a series of vectors combined into a tabular structure. Dataframes have a particular syntax, which is an extension of vector syntax. First let's use the $ sign to call a column

```{r}
head(iris)

colnames(iris)

#How do we call a specific column?
x<-iris$Sepal.Length
```

We can use the same syntax as for vectors with one addiiton.  Indexing a dataframe is given by nameofdf[rows,columns], which can either be the name of the rows or columns, or the column number and row
```{r}
colnames(iris)

#Identical to iris$Sepal.Length

iris[,"Sepal.Length"]
iris[,1]

#dataframes can be indexed for both rows and columns
#remember that simple ranges can be given
1:5

#Get the first five rows and all columns
iris[1:5,]

#Get all rows and the first two columns
a<-iris[,1:2]
head(a)
```

**Try It!**
------------
5. What is the 9th entry of the Sepal.Width column?
6. How would you get all entries for the 17th row?
7. Return an object with the 1 4 and 7 rows of the dataframe?
8. Use the seq command to get all odd rows in the dataset?
9. What happens when you use negative numbers?

Using logical statements
------------------------

An extremely helpful tool is to subset your data using logic rather than an index. Let's begin with continious numeric data.


```{r}
#Grab the Sepal.Width
x<-iris$Petal.Width

#Historgram
hist(x)

#Which values are greater than 1?
logi<-x > 1

#Return all columns based on iris$Petal.Width
a<-iris[logi,]

head(a)

#Return the same column as was subsetted.
x[logi]

#Return a different column based on the subset, ie. Return all values in Sepal.Length where Sepal.Width is greater than 1. 
iris[logi,"Petal.Length"]
```


**Try It!**
------------
10. Why is iris[iris>3,] a nonsensical command? 
11. What about iris[iris$Sepal.Length >3]? 
12. Create a histogram of petal lengths for the entire data
13. Subset the data for values greater than two
14. Create a histogram of your new data


Other types of logical statements 
---------------------------------

Many different types of logical statements can be used to subset data.

For all types of data, if we want a specific value/character/factor we use ==

```{r}
#Get records with petal length of 4
iris[iris$Petal.Length==4,]

#Get only records from the species versicolor
iris[iris$Species=="versicolor",]
```

Logical statments can be combined using & (and) if you want both statements to be true, or | (or) if you want either statement to be true.

```{r}
iris[iris$Petal.Length > 4 & iris$Species == "versicolor",]
```

**Try It!**
------------
Explain in words each of the following logical statements

15. iris[1:4,]

16.  iris[c(1:15,),c(1,3)]

17. iris[iris$Species == "setosa","Petal.Width"]

18. What happens when you add a ! before a logical statment, 
*Hint* compareiris[iris$Species == "setosa",] with iris[!iris$Species == "setosa",]
