
Subsetting Data
=================

R has many powerful subset operators and mastering them will allow you to easily perform complex operation on any kind of dataset. Allows you to manipulate data very succinctly.

Subsetting atomic vectors
```{r}
x <- seq(0,10,1)
```

1. Using positive integers

```{r}
x[1]
x[c(3,1)]
```

2. Using negative integers

```{r}
# skip the first element
x[-1]
# skip the first and the fifth
x[-c(1,5)]
```

3. Using logical operators

```{r}
x[c(TRUE, TRUE, FALSE, FALSE)]
# or based on a condition
x[x > 3]
x[which(x >3)]
Also see `which.max()` and `which.min()`
```

Beyond Vectors - Beginning with Data frames
===========================================

A data frame is a very important data type in R. It's pretty much the de facto data structure for most tabular data and what we use for statistics. Data frames can have additional attributes such as `rownames()` and 'colnames()'. This can be useful for annotating data.

Data frames are usually read in from file, but R comes with 
**Useful functions**  
`head()` - see first 5 rows
`tail()` - see last 5 rows
`dim()` - see dimensions
`nrow()` - number of rows
`ncol()` - number of columns
`str()` - structure of each column
`names()` - will list column names for a data.frame (or any object really).

A data frame is a special type of list where every element of a list has same length.

See that it is actually a special list:

    > is.list(iris)
    [1] TRUE
    > class(iris)
    [1] "data.frame"
     > 
--

**Naming objects**  

Other R objects can also have names not just true for data.frames. Adding names is helpful since it's useful for readable code and self describing objects.

```
x <- 1:3
names(x) <- c("karthik", "ram", "rocks")
x
```


---


# Subsetting data frames

```coffee
df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3])

df[df$x == 2, ]
#   x y z
# 2 2 2 b
df[c(1, 3), ]
#   x y z
# 1 1 3 a
# 3 3 1 c

# There are two ways to select columns from a data frame
# Like a list:
df[c("x", "z")]
#   x z
# 1 1 a
# 2 2 b
# 3 3 c
# Like a matrix
df[, c("x", "z")]
#   x z
# 1 1 a
# 2 2 b
# 3 3 c

# There's an important difference if you select a simple column:
  # matrix subsetting simplifies by default, list subsetting does not.
  df["x"]
#   x
# 1 1
# 2 2
# 3 3
df[, "x"]
# [1] 1 2 3
```
