Worksheet 5 - Data Visualization
========================================================

Let's recall the iris dataset

```{r, some_data, tidy=FALSE, echo=TRUE}
head(iris)

```

```{r,plyrexample , fig.width=6, fig.height=4, out.width='.75\\linewidth', fig.show='hold',  tidy=FALSE}
iris[1:2, ]
# Note the use of the '.' function to allow 'Species' to be used 
# without quoting
ddply(iris, .(Species), summarize, 
      mean.Sep.Wid = mean(Sepal.Width, na.rm = TRUE))
```


```{r,reshapeexample , fig.width=6, fig.height=4, out.width='.75\\linewidth', fig.show='hold',  tidy=FALSE}

iris[1:2, ]
df  <- melt(iris, id.vars = "Species")
df[1:2, ]
```

```{r,reshapeexample2 , fig.width=6, fig.height=4, out.width='.75\\linewidth', fig.show='hold',  tidy=FALSE}
df[1:2, ]
dcast(df, Species ~ variable, mean)
```

```{r,data_summary, echo=TRUE,  tidy=FALSE}
myplot <- ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width))
summary(myplot)
```

```{r,first_plotb , eval=FALSE, tidy=FALSE}
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width))
+ geom_point()

myplot <- ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width))
myplot + geom_point()
```

## ----first_plot , fig.width=6, fig.height=4, out.width='.75\\linewidth', fig.show='hold', fig.align='center',  tidy=FALSE----
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width)) +
geom_point()


## ----first_plot_size , fig.width=6, fig.height=4, out.width='.75\\linewidth', fig.show='hold', fig.align='center',  tidy=FALSE----
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width)) +
geom_point(size = 3)


## ----first_plot_color , fig.width=6, fig.height=4, out.width='.75\\linewidth', fig.show='hold', fig.align='center',  tidy=FALSE----
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
geom_point(size = 3)


## ----first_plot_shape , fig.width=6, fig.height=3.5, out.width='.75\\linewidth', fig.show='hold', fig.align='center',  tidy=FALSE----
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
geom_point(aes(shape = Species), size = 3)
# Why aes(shape = Species)?
